---
title: "Simulation study: CLMM vs LMEM for 5-category ordinal outcomes"
author: "Your Name Here"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    number-sections: true
    theme: cosmo
execute:
  warning: false
  message: false
---

```{r setup}
#| label: setup
#| include: true

# ---- Package loading -------------------------------------------------------
# Core modelling packages
library(lme4)      # LMEM: linear mixed-effects models
library(ordinal)   # CLMM: cumulative link mixed models (clmm())

# Data wrangling / tidyverse
library(tidyverse) # dplyr, tidyr, purrr, ggplot2, tibble, etc.
library(broom.mixed) # tidy() extraction for mixed models

# Optional: parallel simulation (uncomment if using furrr instead of for-loops)
# library(furrr)
# library(future)

# Tables
library(knitr)
library(kableExtra)

# Reproducibility - see USER_SEED in Simulation Parameters section below
```

## Abstract

Ordinal outcomes with a small number of categories (e.g., 5-point Likert-type
ratings) are common in psychology and cognitive science, yet they are
routinely analysed with linear mixed-effects models (LMEM) that treat
category scores as continuous, interval-scaled data. Cumulative link mixed
models (CLMM) explicitly respect the ordinal, bounded nature of such
outcomes, but it is unclear how much this matters in realistic
repeated-measures designs with multiple between-subjects Likert-scored
predictors. This simulation study, structured according to the ADEMP
framework (Aims, Data-generating mechanisms, Estimands, Methods, Performance
measures; Morris, White, & Crowther, 2019), compares CLMM and LMEM for a
5-category ordinal outcome with three between-subjects predictors (Group1,
Group2, Group3) and one within-subject factor (Condition). Data are
generated from a true cumulative logit mixed model containing only main
effects, and both CLMM and LMEM are fitted with a Group x Condition
interaction term to allow evaluation of Type I error alongside power. We
vary distributional shape, intraclass correlation (ICC), effect size, and
sample size. Performance is assessed via bias, empirical standard deviation,
mean squared error, Type I error, power, and confidence interval coverage.
**Results are not yet available and will be inserted by the user after
running the simulation locally.**

## Introduction

Likert-type and other ordered-categorical response scales pervade
psychological and cognitive science research: confidence ratings, agreement
scales, symptom severity scores, and many experimental judgement tasks
produce outcomes with a small, fixed number of ordered categories (often 5
or 7). Despite the ordinal nature of these data, it remains standard
practice to code categories as consecutive integers and to analyse them
with linear (mixed-effects) models, implicitly assuming equal spacing
between categories and normally distributed, homoscedastic residuals
(Liddell & Kruschke, 2018). This "metric" treatment of ordinal data can
distort estimated effect sizes, inflate or deflate Type I error rates, and
produce biased inferences, particularly when response distributions are
skewed, polarised, or bounded near the scale limits (Bürkner & Vuorre,
2019; Liddell & Kruschke, 2018).

Cumulative link (mixed) models (CLMMs) provide a principled alternative:
they model the probability of being in category k or below as a function
of latent thresholds and a linear predictor, without assuming equal
inter-category distances (McCullagh, 1980; Agresti, 2010). Extending this
to hierarchical/repeated-measures designs, CLMMs incorporate participant-
level random effects in the same way LMEMs do, but on the latent (logit or
probit) scale rather than the observed integer scale (Christensen, 2022,
`ordinal` package). Foundational and applied work on ordinal versus normal
mixed models (e.g., Hedeker & Gibbons, 1994; Hedeker, 2008; Johnson, 2003)
has long argued that treating ordinal data as continuous can bias
inference, but direct, systematic comparisons in designs typical of
psychology — repeated within-subject conditions crossed with *multiple*
continuous/Likert-scored between-subjects predictors — remain relatively
scarce.

This study addresses that gap using a Monte Carlo simulation designed
according to the ADEMP framework (Morris, White, & Crowther, 2019), which
structures simulation studies around explicit statements of Aims,
Data-generating mechanisms, Estimands, Methods, and Performance measures.
We simulate 5-category ordinal outcomes from a true cumulative logit mixed
model with three between-subjects Likert-scored predictors and one
within-subject factor, and compare CLMM and LMEM fits — each including a
Group x Condition interaction term whose true value is zero — across
variations in distributional shape, ICC, effect size, and sample size.

## Aims

- **Aim 1:** How do CLMM and LMEM compare in terms of Type I error and
  power for detecting Group main effects and the Group x Condition
  interaction (whose true value is zero in the data-generating model)?
- **Aim 2:** How does this comparison depend on distributional shape
  (symmetric vs skewed vs polarised), intraclass correlation (ICC), and
  sample size?
- **Aim 3:** Does CLMM provide less biased estimates of fixed effects than
  LMEM when the true outcome is ordinal with 5 categories?

## Methods (ADEMP)

The simulation follows the ADEMP structure of Morris et al. (2019).
Sub-sections below correspond to Data-generating mechanisms, Estimands,
Methods of analysis, and Performance measures. All user-editable simulation
values are isolated in a single chunk (@sec-user-params) so nothing below
needs hard-coded numbers changed by hand.

### Data-generating mechanisms (D)

Data are generated from a cumulative logit (or probit) mixed model:

$$
\text{logit}\,P(Y_{it} \le k) = \alpha_k - \big(\beta_0 + \beta_1 \text{Group1}_i + \beta_2 \text{Group2}_i + \beta_3 \text{Group3}_i + \beta_4 \text{Condition}_t + b_{0i} + b_{1i}\text{Condition}_t\big)
$$

for $k = 1, \dots, 4$ thresholds separating $K = 5$ ordered categories, with
$b_{0i}, b_{1i} \sim N(0, \Sigma)$ jointly normal participant-level random
intercept and slope for Condition. Group1, Group2, and Group3 are
participant-level, between-subjects predictors, each generated on a 1-5
Likert scale and (optionally) centered/standardised before entering the
linear predictor. Condition is a within-subject factor with multiple
levels. **No Group x Condition interaction terms are included in the true
DGM** — all true interaction coefficients are zero — but both analysis
models below still include a Group1:Condition interaction term so that its
Type I error rate can be evaluated directly.

We vary:

- **Effect size** for Group and Condition main effects (null vs small vs
  medium vs large, on the latent logit/probit scale).
- **Distributional shape**, via the threshold set $\alpha_1 < \dots <
  \alpha_4$: symmetric, skewed, and polarised (extreme-heavy) category
  probability profiles.
- **ICC**, via the random-intercept/slope variances and their correlation
  (low, medium, high).
- **Sample size** ($N$ participants), at least two levels.

The true interaction coefficient is fixed at zero across *all* scenarios.

### Estimands (E)

- **Primary:** $\beta_1, \beta_2, \beta_3$ (Group1/2/3 main effects on the
  latent scale), and the Group1:Condition interaction coefficient fitted by
  each analysis model (true value = 0).
- **Secondary:** $\beta_4$ (Condition main effect), thresholds $\alpha_k$,
  random-effect variances/correlation, and implied ICC.

### Methods of analysis (M)

For every simulated dataset we fit:

- **CLMM:** `ordinal::clmm(Y ~ Group1 + Group2 + Group3 + Condition + Group1:Condition + (1 + Condition | id), link = "logit")`
  (switch `link = "probit"` to compare link functions). Wald z-tests are
  used for fixed-effect p-values (the default reported by `clmm`); an LRT
  alternative via `anova()` on nested models is noted in comments for users
  who prefer it.
- **LMEM:** `lme4::lmer(Ynum ~ Group1 + Group2 + Group3 + Condition + Group1:Condition + (1 + Condition | id))` on `Ynum = as.numeric(Y)`,
  with Wald (or Satterthwaite/`lmerTest`, noted in comments) p-values.

### Performance measures (P)

For each estimand, method, and scenario, across simulation repetitions we
compute: bias, empirical SD, MSE, Type I error (for null effects, including
the interaction), power (for non-null main effects), and 95% CI coverage,
each with Monte Carlo standard errors (MCSE), following the formulae in
Morris et al. (2019, Table 6).

---

## Simulation Parameters (User-Specified) {#sec-user-params}

```{r sim-params}
#| label: sim-params

# =============================================================================
# SIMULATION PARAMETERS -- USER SPECIFIED
# Edit the values below to match your own study. All downstream code
# references these objects; do NOT hard-code numbers elsewhere.
# =============================================================================

## --- Reproducibility -------------------------------------------------------
USER_SEED <- 20260819L  # random seed; change to your own preferred seed

## --- Sample sizes -----------------------------------------------------------
# Number of participants per scenario. Provide >= 2 levels to study how
# performance scales with N.
N_LEVELS <- c(100, 200)

## --- Within-subject Condition factor ----------------------------------------
N_CONDITION_LEVELS <- 4          # number of within-subject Condition levels
# Coding/spacing of Condition in the linear predictor. If TRUE, Condition is
# treated as a numeric/ordered variable with equally spaced scores
# (e.g., -1.5, -0.5, 0.5, 1.5 for 4 levels); if FALSE, Condition is entered
# as a factor with (e.g.) treatment/contrast coding -- edit CONDITION_CONTRASTS.
CONDITION_NUMERIC <- TRUE
CONDITION_SCORES  <- scale(seq_len(N_CONDITION_LEVELS), center = TRUE, scale = FALSE)[, 1]
# If CONDITION_NUMERIC == FALSE, specify a contrast matrix, e.g.:
# CONDITION_CONTRASTS <- contr.sum(N_CONDITION_LEVELS)

## --- Monte Carlo repetitions -------------------------------------------------
N_SIM_REPS <- 500   # simulation repetitions per scenario (increase to 1000+ for final runs)

## --- True fixed-effect values (latent logit/probit scale) -------------------
# Group1/Group2/Group3 main effects: null value (for Type I error) plus
# small/medium/large non-null values (for power). Edit magnitudes to reflect
# effect sizes plausible for your substantive area (these are placeholder
# defaults on the latent logistic scale, loosely following Cohen-like
# conventions rescaled for cumulative logit models).
BETA_GROUP_NULL   <- 0.00
BETA_GROUP_SMALL  <- 0.20
BETA_GROUP_MEDIUM <- 0.50
BETA_GROUP_LARGE  <- 0.80
BETA_GROUP_LEVELS <- c(null = BETA_GROUP_NULL, small = BETA_GROUP_SMALL,
                       medium = BETA_GROUP_MEDIUM, large = BETA_GROUP_LARGE)

# True Condition main effect (latent scale). Edit as needed; can also be
# varied across a vector of levels analogous to BETA_GROUP_LEVELS.
BETA_CONDITION <- 0.40

# True Group x Condition interaction is ALWAYS zero in the DGM (do not edit
# to a non-zero value -- this is required for the Type I error evaluation
# described in the Aims).
BETA_INTERACTION_TRUE <- 0.00

## --- Threshold sets (alpha_1..alpha_4) for K = 5 categories -----------------
# Each vector must be strictly increasing (4 thresholds for 5 categories).
# Symmetric: roughly equal mass across categories.
THRESH_SYMMETRIC <- c(-1.5, -0.5, 0.5, 1.5)
# Skewed: more mass concentrated in lower categories (positive skew in
# category counts); reverse signs/order for negative skew.
THRESH_SKEWED    <- c(-2.5, -1.0, 0.2, 1.2)
# Polarized: more mass at the extreme categories, little in the middle.
THRESH_POLARIZED <- c(-2.0, -0.3, 0.3, 2.0)

THRESHOLD_SETS <- list(
  symmetric = THRESH_SYMMETRIC,
  skewed    = THRESH_SKEWED,
  polarized = THRESH_POLARIZED
)

## --- Random-effects structure (participant intercept + Condition slope) -----
# SDs and correlation for low/medium/high ICC scenarios. Increasing the
# random-intercept SD relative to residual/latent-scale variance increases
# ICC; edit magnitudes to bracket the range plausible for your paradigm.
RANEF_SETTINGS <- tibble::tribble(
  ~icc_level, ~sd_intercept, ~sd_slope, ~cor_int_slope,
  "low",       0.5,           0.2,       0.0,
  "medium",    1.0,           0.3,       0.1,
  "high",      1.8,           0.4,       0.2
)

## --- Between-subjects Likert predictors (Group1/Group2/Group3) -------------
# Distribution used to generate each Group predictor on a 1-5 Likert scale.
# Options implemented below: "discrete_uniform" (equal probability 1-5) or
# "rounded_normal" (normal draw rounded and clipped to [1, 5]).
GROUP_GEN_METHOD <- "rounded_normal"   # "discrete_uniform" or "rounded_normal"
GROUP_NORMAL_MEAN <- 3      # midpoint of 1-5 scale, used if rounded_normal
GROUP_NORMAL_SD   <- 1      # spread, used if rounded_normal

# Whether to center (mean 0) and/or standardize (SD 1) Group1/2/3 before
# they enter the linear predictor. Centering is recommended at minimum so
# that beta_0 / thresholds retain a sensible reference (average participant).
GROUP_CENTER    <- TRUE
GROUP_STANDARDIZE <- FALSE

## --- Link function ----------------------------------------------------------
LINK_FUNCTION <- "logit"   # "logit" or "probit"; used both in DGM and in clmm()

## --- Derived / convenience objects (do not need editing) --------------------
set.seed(USER_SEED)
```

### Helper functions

```{r helper-functions}
#| label: helper-functions

#' Generate a 1-5 Likert-scale between-subjects predictor
#'
#' @param n number of participants
#' @param method "discrete_uniform" or "rounded_normal"
#' @param mean_ mean for rounded_normal method
#' @param sd_ sd for rounded_normal method
#' @return integer vector length n, values in 1:5
generate_likert_predictor <- function(n,
                                       method = GROUP_GEN_METHOD,
                                       mean_ = GROUP_NORMAL_MEAN,
                                       sd_ = GROUP_NORMAL_SD) {
  if (method == "discrete_uniform") {
    sample(1:5, size = n, replace = TRUE)
  } else if (method == "rounded_normal") {
    x <- round(rnorm(n, mean = mean_, sd = sd_))
    pmin(pmax(x, 1), 5)
  } else {
    stop("Unknown method for generate_likert_predictor()")
  }
}

#' Rescale a raw 1-5 predictor according to user preferences (center/standardize)
rescale_predictor <- function(x, center = GROUP_CENTER, standardize = GROUP_STANDARDIZE) {
  out <- x
  if (center) out <- out - mean(out)
  if (standardize) out <- out / sd(out)
  out
}

#' Simulate one dataset from the true CLMM data-generating model
#'
#' @param n_participants number of participants (N)
#' @param n_condition number of within-subject Condition levels
#' @param condition_scores numeric vector of Condition scores (length n_condition)
#' @param thresholds numeric vector of K-1 increasing thresholds
#' @param beta_group numeric vector length 3: true Group1/2/3 coefficients
#' @param beta_condition true Condition main-effect coefficient
#' @param beta_interaction true Group1:Condition interaction coefficient
#'   (should be 0 for all scenarios per study design)
#' @param sd_intercept random-intercept SD
#' @param sd_slope random Condition-slope SD
#' @param cor_int_slope correlation between random intercept and slope
#' @param link "logit" or "probit"
#' @return tibble, one row per observation (long format), columns:
#'   id, Group1, Group2, Group3, Condition (factor), Condition_num, Y (ordered
#'   factor), Ynum (numeric copy of Y)
simulate_clmm_data <- function(n_participants,
                                n_condition = N_CONDITION_LEVELS,
                                condition_scores = CONDITION_SCORES,
                                thresholds = THRESH_SYMMETRIC,
                                beta_group = c(0, 0, 0),
                                beta_condition = BETA_CONDITION,
                                beta_interaction = BETA_INTERACTION_TRUE,
                                sd_intercept = 1,
                                sd_slope = 0.3,
                                cor_int_slope = 0,
                                link = LINK_FUNCTION) {

  K <- length(thresholds) + 1

  # Participant-level between-subjects Likert predictors
  Group1 <- rescale_predictor(generate_likert_predictor(n_participants))
  Group2 <- rescale_predictor(generate_likert_predictor(n_participants))
  Group3 <- rescale_predictor(generate_likert_predictor(n_participants))

  # Correlated random intercept/slope via bivariate normal
  ranef_cov <- matrix(
    c(sd_intercept^2, cor_int_slope * sd_intercept * sd_slope,
      cor_int_slope * sd_intercept * sd_slope, sd_slope^2),
    nrow = 2
  )
  ranef_draws <- MASS::mvrnorm(n_participants, mu = c(0, 0), Sigma = ranef_cov)
  b0 <- ranef_draws[, 1]
  b1 <- ranef_draws[, 2]

  participant_df <- tibble(
    id = seq_len(n_participants),
    Group1 = Group1, Group2 = Group2, Group3 = Group3,
    b0 = b0, b1 = b1
  )

  # Expand to long format: one row per participant x Condition level
  long_df <- participant_df %>%
    tidyr::expand_grid(cond_index = seq_len(n_condition)) %>%
    dplyr::mutate(
      Condition_num = condition_scores[cond_index],
      Condition = factor(cond_index, labels = paste0("C", seq_len(n_condition)))
    ) %>%
    dplyr::arrange(id, cond_index)

  # Linear predictor (eta), NOTE: no true Group:Condition interaction term
  long_df <- long_df %>%
    dplyr::mutate(
      eta = beta_group[1] * Group1 + beta_group[2] * Group2 + beta_group[3] * Group3 +
        beta_condition * Condition_num +
        beta_interaction * Group1 * Condition_num +  # kept for completeness; should be 0
        b0 + b1 * Condition_num
    )

  # Convert linear predictor into category probabilities via cumulative link
  link_fun    <- if (link == "probit") pnorm else plogis
  cum_probs <- sapply(thresholds, function(a) link_fun(a - long_df$eta))
  cum_probs <- cbind(cum_probs, 1)  # P(Y <= K) = 1
  cat_probs <- cbind(cum_probs[, 1], t(diff(t(cum_probs))))
  cat_probs[cat_probs < 0] <- 0
  cat_probs <- cat_probs / rowSums(cat_probs)

  y_draw <- apply(cat_probs, 1, function(p) sample.int(K, size = 1, prob = p))

  long_df %>%
    dplyr::mutate(
      Y = factor(y_draw, levels = 1:K, ordered = TRUE),
      Ynum = as.numeric(y_draw)
    ) %>%
    dplyr::select(id, Group1, Group2, Group3, Condition, Condition_num, Y, Ynum)
}

#' Fit the CLMM analysis model to one simulated dataset
#'
#' Analysis model always includes the Group1:Condition interaction, even
#' though the true DGM has no interaction, so that Type I error for the
#' interaction term can be assessed. Fixed-effect inference uses Wald
#' z-tests as returned by summary(clmm_fit); see comment below for an LRT
#' alternative.
fit_clmm <- function(dat, link = LINK_FUNCTION) {
  fit <- tryCatch(
    ordinal::clmm(
      Y ~ Group1 + Group2 + Group3 + Condition_num + Group1:Condition_num +
        (1 + Condition_num | id),
      data = dat, link = link
    ),
    error = function(e) NULL
  )
  if (is.null(fit)) return(NULL)

  # Wald-based tidy summary (default). For an LRT alternative on a specific
  # term, fit a reduced model without that term and compare via anova():
  #   fit_reduced <- update(fit, . ~ . - Group1:Condition_num)
  #   anova(fit_reduced, fit)
  broom.mixed::tidy(fit, effects = "fixed") %>%
    dplyr::mutate(method = "CLMM")
}

#' Fit the LMEM analysis model to the numeric-coded version of the same data
fit_lmem <- function(dat) {
  fit <- tryCatch(
    lme4::lmer(
      Ynum ~ Group1 + Group2 + Group3 + Condition_num + Group1:Condition_num +
        (1 + Condition_num | id),
      data = dat, REML = TRUE
    ),
    error = function(e) NULL
  )
  if (is.null(fit)) return(NULL)

  # Wald p-values from lme4 by default (no df correction). For
  # Satterthwaite/Kenward-Roger df, refit with lmerTest::lmer() instead.
  broom.mixed::tidy(fit, effects = "fixed") %>%
    dplyr::mutate(
      p.value = 2 * pnorm(abs(statistic), lower.tail = FALSE),  # Wald z-approx
      method = "LMEM"
    )
}
```

### Scenario grid

```{r scenario-grid}
#| label: scenario-grid

# One row per unique combination of design factors. Each row is later
# expanded into N_SIM_REPS simulated datasets.
scenario_grid <- tidyr::expand_grid(
  N               = N_LEVELS,
  shape           = names(THRESHOLD_SETS),
  icc_level       = RANEF_SETTINGS$icc_level,
  group_effect    = names(BETA_GROUP_LEVELS)
) %>%
  dplyr::left_join(RANEF_SETTINGS, by = "icc_level") %>%
  dplyr::mutate(
    scenario_id  = dplyr::row_number(),
    beta_group1  = BETA_GROUP_LEVELS[group_effect],
    beta_group2  = BETA_GROUP_LEVELS[group_effect],  # edit if Group2/3 should differ
    beta_group3  = BETA_GROUP_LEVELS[group_effect],
    thresholds   = THRESHOLD_SETS[shape]
  )

# Inspect the grid before running the full simulation:
# print(scenario_grid, n = Inf)
```

### Simulation run

```{r simulation-run}
#| label: simulation-run
#| eval: false

# -----------------------------------------------------------------------
# NOTE: This chunk can be slow (N_SIM_REPS x nrow(scenario_grid) model
# fits, each involving both a CLMM and an LMEM fit). Set eval: true only
# when ready to run the full simulation. For quicker iteration, reduce
# N_SIM_REPS temporarily, or parallelize with furrr::future_map_dfr()
# (commented alternative below).
# -----------------------------------------------------------------------

run_one_replicate <- function(scenario_row, rep_id) {
  dat <- simulate_clmm_data(
    n_participants   = scenario_row$N,
    thresholds       = scenario_row$thresholds[[1]],
    beta_group       = c(scenario_row$beta_group1, scenario_row$beta_group2, scenario_row$beta_group3),
    beta_condition   = BETA_CONDITION,
    beta_interaction = BETA_INTERACTION_TRUE,
    sd_intercept     = scenario_row$sd_intercept,
    sd_slope         = scenario_row$sd_slope,
    cor_int_slope    = scenario_row$cor_int_slope
  )

  clmm_est <- fit_clmm(dat)
  lmem_est <- fit_lmem(dat)

  dplyr::bind_rows(clmm_est, lmem_est) %>%
    dplyr::mutate(
      scenario_id = scenario_row$scenario_id,
      rep_id = rep_id,
      N = scenario_row$N,
      shape = scenario_row$shape,
      icc_level = scenario_row$icc_level,
      group_effect = scenario_row$group_effect
    )
}

# Simple (sequential) loop version:
sim_results <- purrr::map_dfr(seq_len(nrow(scenario_grid)), function(i) {
  scenario_row <- scenario_grid[i, ]
  purrr::map_dfr(seq_len(N_SIM_REPS), function(r) {
    run_one_replicate(scenario_row, rep_id = r)
  })
})

# Parallelized alternative (uncomment and adapt; requires furrr + future):
# future::plan(future::multisession, workers = parallel::detectCores() - 1)
# sim_results <- furrr::future_map_dfr(
#   seq_len(nrow(scenario_grid)),
#   function(i) {
#     scenario_row <- scenario_grid[i, ]
#     purrr::map_dfr(seq_len(N_SIM_REPS), ~ run_one_replicate(scenario_row, rep_id = .x))
#   },
#   .options = furrr::furrr_options(seed = TRUE)
# )

# Save raw simulation output for later reuse without re-running:
# saveRDS(sim_results, "sim_results.rds")
```

### Performance measure computation

```{r performance-measures}
#| label: performance-measures
#| eval: false

# True values for each estimand, keyed by scenario characteristics, used to
# compute bias / Type I error / power / coverage. Edit mapping if term names
# differ after tidy() (check unique(sim_results$term) first).
true_value_lookup <- function(term, group_effect) {
  dplyr::case_when(
    term == "Group1" ~ BETA_GROUP_LEVELS[group_effect],
    term == "Group2" ~ BETA_GROUP_LEVELS[group_effect],
    term == "Group3" ~ BETA_GROUP_LEVELS[group_effect],
    term == "Condition_num" ~ BETA_CONDITION,
    term == "Group1:Condition_num" ~ BETA_INTERACTION_TRUE,
    TRUE ~ NA_real_
  )
}

perf_summary <- sim_results %>%
  dplyr::filter(term %in% c("Group1", "Group2", "Group3",
                             "Condition_num", "Group1:Condition_num")) %>%
  dplyr::mutate(
    true_value = true_value_lookup(term, group_effect),
    is_null_true = near(true_value, 0),
    ci_lower = estimate - 1.96 * std.error,
    ci_upper = estimate + 1.96 * std.error,
    covered  = true_value >= ci_lower & true_value <= ci_upper,
    rejected = p.value < 0.05
  ) %>%
  dplyr::group_by(method, term, N, shape, icc_level, group_effect, true_value) %>%
  dplyr::summarise(
    n_reps      = dplyr::n(),
    bias        = mean(estimate - true_value, na.rm = TRUE),
    bias_mcse   = sd(estimate - true_value, na.rm = TRUE) / sqrt(n_reps),
    emp_sd      = sd(estimate, na.rm = TRUE),
    mse         = mean((estimate - true_value)^2, na.rm = TRUE),
    mse_mcse    = sd((estimate - true_value)^2, na.rm = TRUE) / sqrt(n_reps),
    reject_rate = mean(rejected, na.rm = TRUE),           # Type I error if true_value == 0, else power
    reject_mcse = sqrt(reject_rate * (1 - reject_rate) / n_reps),
    coverage    = mean(covered, na.rm = TRUE),
    coverage_mcse = sqrt(coverage * (1 - coverage) / n_reps),
    .groups = "drop"
  )

# saveRDS(perf_summary, "perf_summary.rds")
```

---

## Results

*Placeholder: this section intentionally contains no results yet. After
running the simulation chunks above (`simulation-run` and
`performance-measures`) locally, populate `sim_results` and `perf_summary`,
then set `eval: true` on the chunks below to generate the tables and plots.*

```{r results-table-bias}
#| label: results-table-bias
#| eval: false

perf_summary %>%
  dplyr::filter(term %in% c("Group1", "Group2", "Group3")) %>%
  dplyr::select(method, term, N, shape, icc_level, group_effect,
                bias, bias_mcse, emp_sd, mse, mse_mcse) %>%
  knitr::kable(digits = 3, caption = "Bias, empirical SD, and MSE for Group main effects, by method and scenario.") %>%
  kableExtra::kable_styling(full_width = FALSE)
```

```{r results-table-typeI-power}
#| label: results-table-typeI-power
#| eval: false

perf_summary %>%
  dplyr::select(method, term, N, shape, icc_level, group_effect,
                true_value, reject_rate, reject_mcse, coverage, coverage_mcse) %>%
  knitr::kable(digits = 3, caption = "Type I error / power (reject_rate) and 95% CI coverage, by method, term, and scenario.") %>%
  kableExtra::kable_styling(full_width = FALSE)
```

```{r results-plot-power-icc}
#| label: results-plot-power-icc
#| eval: false

perf_summary %>%
  dplyr::filter(term == "Group1") %>%
  ggplot2::ggplot(ggplot2::aes(x = icc_level, y = reject_rate, colour = method, group = method)) +
  ggplot2::geom_point(position = ggplot2::position_dodge(width = 0.2)) +
  ggplot2::geom_line(ggplot2::aes(group = interaction(method)), position = ggplot2::position_dodge(width = 0.2)) +
  ggplot2::facet_grid(shape ~ group_effect) +
  ggplot2::labs(
    x = "ICC level", y = "Rejection rate (Type I error / power)",
    colour = "Method",
    title = "Rejection rate for Group1 effect across ICC, shape, and effect-size scenarios"
  ) +
  ggplot2::theme_minimal()
```

```{r results-plot-bias-shape}
#| label: results-plot-bias-shape
#| eval: false

perf_summary %>%
  dplyr::filter(term %in% c("Group1", "Group2", "Group3")) %>%
  ggplot2::ggplot(ggplot2::aes(x = shape, y = bias, colour = method, group = method)) +
  ggplot2::geom_point(position = ggplot2::position_dodge(width = 0.2)) +
  ggplot2::geom_hline(yintercept = 0, linetype = "dashed") +
  ggplot2::facet_grid(term ~ N, labeller = ggplot2::label_both) +
  ggplot2::labs(
    x = "Distribution shape", y = "Bias",
    colour = "Method",
    title = "Bias of Group main-effect estimates across distributional shapes and sample sizes"
  ) +
  ggplot2::theme_minimal()
```

```{r results-plot-interaction-typeI}
#| label: results-plot-interaction-typeI
#| eval: false

perf_summary %>%
  dplyr::filter(term == "Group1:Condition_num") %>%
  ggplot2::ggplot(ggplot2::aes(x = icc_level, y = reject_rate, colour = method, group = method)) +
  ggplot2::geom_point(position = ggplot2::position_dodge(width = 0.2)) +
  ggplot2::geom_hline(yintercept = 0.05, linetype = "dashed") +
  ggplot2::facet_wrap(~ shape) +
  ggplot2::labs(
    x = "ICC level", y = "Type I error rate (interaction term)",
    colour = "Method",
    title = "Type I error rate for the Group1:Condition interaction (true value = 0)"
  ) +
  ggplot2::theme_minimal()
```

## Discussion

*This section is a generic outline only; specific conclusions should be
added once results are available.*

- **When CLMM would be preferred:** CLMMs are expected to show advantages
  over LMEMs particularly when the outcome distribution is skewed or
  polarised rather than symmetric, when the number of categories is small
  (as here, K = 5), and when effect sizes on the latent scale are
  moderate-to-large. Discuss whether the simulation supports this pattern.
- **When the two methods converge:** Under symmetric distributions with
  many categories and low ICC, LMEM often approximates CLMM reasonably
  well; discuss whether this held here, and for which estimands (Group
  main effects vs the interaction Type I error) the two methods diverged
  most.
- **Practical recommendations:** Provide guidance for applied researchers
  on choosing between CLMM and LMEM for 5-point Likert outcomes in
  repeated-measures designs with multiple between-subjects Likert-scored
  predictors, referencing the specific bias/power/Type I error patterns
  found.
- **Limitations:** Note constraints such as the number of scenarios
  simulated, the specific random-effects structure assumed (random
  intercept + Condition slope only), the use of Wald-based inference for
  both models, and the restriction to a single true DGM form (cumulative
  logit with only main effects).
- **Future work:** Extensions could include non-cumulative-logit DGMs
  (e.g., adjacent-category or continuation-ratio models), true non-zero
  interaction effects, model misspecification scenarios (e.g., fitting
  CLMM with the wrong link), and comparison against Bayesian ordinal models
  (e.g., via `brms`).

## References

Agresti, A. (2010). *Analysis of Ordinal Categorical Data* (2nd ed.). Wiley.

Bürkner, P.-C., & Vuorre, M. (2019). Ordinal regression models in
psychology: A tutorial. *Advances in Methods and Practices in Psychological
Science*, 2(1), 77-101.

Christensen, R. H. B. (2022). *ordinal: Regression Models for Ordinal Data*
[R package]. Retrieved from https://CRAN.R-project.org/package=ordinal

Hedeker, D. (2008). Multilevel models for ordinal and nominal variables. In
J. de Leeuw & E. Meijer (Eds.), *Handbook of Multilevel Analysis* (pp.
237-274). Springer.

Hedeker, D., & Gibbons, R. D. (1994). A random-effects ordinal regression
model for multilevel analysis. *Biometrics*, 50(4), 933-944.

Johnson, T. R. (2003). On the use of heterogeneous thresholds ordinal
regression models to account for individual differences in response style.
*Psychometrika*, 68(4), 563-583.

Liddell, T. M., & Kruschke, J. K. (2018). Analyzing ordinal data with
metric models: What could possibly go wrong? *Journal of Experimental
Social Psychology*, 79, 328-348.

McCullagh, P. (1980). Regression models for ordinal data. *Journal of the
Royal Statistical Society: Series B*, 42(2), 109-127.

Morris, T. P., White, I. R., & Crowther, M. J. (2019). Using simulation
studies to evaluate statistical methods. *Statistics in Medicine*, 38(11),
2074-2102.
